Qualitative Assessesment of Graph Neural Network Potentials for Phonon Properties of Shape Memory Alloys using Graph Neural Network Potentials

### 2024-01-08 src/tex/ms.tex
Some adjustments

\title{Assessing Phonon Properties of Shape Memory Alloys using Graph Neural Network Potentials}

@@

The training of these early GNN utlized large density functional theory calculation repositories and mostly tuned the learning cost function to ground-state structure and energy. This resulted in mean-absolute-errors and other measures of performance that could be improved upon \cite{Riebesell2023}. More recently there have been significant improvements in both training data used and the GNN architecture that show promise for truly approaching a general interatomic potential suitable for static and dynamic atomistic calculations of multicomponents and multiphases. If succesfull, identifying and designing new potential SMA materials using GNN interatomic potentials could have tremendous impact on technology. Therefore, the primary goal is to provide a qualitative assessment of predictions of the equations of state, phonon bandstructure, and elastic constants for various phases of NiTi and PtTi using three GNN potentials intended for general use: M3GNet, CHGNet, and MACE \cite{Chen2022,Deng2023,Batatia2022}. We note that the \citet{Riebesell2023} provides a summarized performance over these GNN using the materials project database, but does not specifically address SMA and phonon characteristics. \par

@@

@@

\begin{figure}[ht!]
    \script{PlotCohesiveEnergy.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/NiTi_CohesiveEnergyPlot.png}
        \caption{
          Comparison of predicted cohesive energy for respective ground-state structure and model.
        }
        \label{fig:ecoh}
    \end{centering}
\end{figure}

>>>>

\title{Qualitative Assessesment of Graph Neural Network Potentials for  Phonon Properties of Shape Memory Alloys using Graph Neural Network Potentials}

@@

The training of these early GNN utlized large density functional theory calculation repositories and mostly tuned the learning cost function to ground-state structure and energy. This resulted in mean-absolute-errors and other measures of performance that could be improved upon \cite{Riebesell2023}. More recently there have been significant improvements in both training data used and the GNN architecture that show promise for truly approaching a general interatomic potential suitable for static and dynamic atomistic calculations of multicomponents and multiphases. If succesfull, identifying and designing new potential SMA materials using GNN interatomic potentials could have tremendous impact on technology. Therefore, the primary goal is to provide a qualitative assessment of pre-trained GNN interatomic potentials that are promoted as being transferable across compositions and phases. We specfically focus on the predictions for equations of state, phonon bandstructure, and elastic constants for various phases of NiTi and PtTi using three GNN potentials: M3GNet, CHGNet, and MACE \cite{Chen2022,Deng2023,Batatia2022}. We note that the \citet{Riebesell2023} provides a summarized performance over these GNN using the materials project database, but does not specifically address SMA and phonon characteristics. \par

@@

One way to understand the phase stability for a given composition is to compare the cohesive energy, which is defined as,

\begin{equation}
  \label{eq:cohesive_energy}
  E_{c} = \frac{E_{\text{total}}}{N}
\end{equation}

With $E_{\text{total}}$ the total energy of the system and $N$ the number of atoms. In fig.~\ref{fig:ecoh} the bar chart for different phases and interatomic potentials models are show. We compare these to the predicted cohesive energy for The Materials Project (MP) entry \texttt{MP-571} which corresponds to the \textit{B2} phase \cite{MP--571}. In the figure we see that the EAM and MEAM potentials show significantly lower cohesive energy compared to the GNN potentials. The stark increase in binding energy for the GNN is surprising as the EAM and MEAM potentials are known to reproduce previous experimental and DFT data \cite{Haskins2016,Ko2015}. The GNNs predict a more strongly bound and stable \textit{B2} structure but all are consistent with the entry for The MP and calculation details can be found in ref.~\cite{MP--571}. \par

\begin{figure}[ht!]
    \script{PlotCohesiveEnergy.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/NiTi_CohesiveEnergyPlot.png}
        \caption{
          Comparison of predicted cohesive energy for respective ground-state structure and model. As a reference we use The Materials Project entry for the \ce{NiTi} \textit{B2} entry \cite{MP--571}.
        }
        \label{fig:ecoh}
    \end{centering}
\end{figure}

@@
